Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-23 EconGrowth.Rnw
Added a comment where Caleb should start adding text.
Cleaned up nomenclature for functions. Functionized CES curve fits. Tried CES curve fit for ZA, but it fails.

@@

<<Load Data, include=FALSE>>=
#################################
# This function loads data given a country name.
#
# countryName = a string of the 2-digit code for the country for which data is to be loaded. 
#   The file name from which data will be loaded is assumed to be of the form 
#   "<countryName>Data.txt". The file is assumed to exist in a subfolder of this project called "data".
#
# returns a matrix with the data that has been loaded
##
loadData <- function(countryName){
  # Read the data file as a table with a header.  
  fileName <- paste("data/", countryName, "Data.txt", sep="")
  dataTable <- read.table(fileName, header = TRUE)
  # Identifies the header names associated with dataTable
  #print(names(dataTable))
  return(dataTable)
}
@

@@

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){

@@

####################################
# Creates a LaTeX printable table from the data. This function first calls cobbDouglasData.
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a printable LaTeX table from xtable.
##
cobbDouglasTable <- function(countryName){
  dataCD <- cobbDouglasData(countryName) 
  colnames(dataCD) <- c("$\\lambda$", "$\\alpha$", "$\\beta$", "$\\gamma$")
  tableCD <- xtable(dataCD, caption=paste(countryName, " Cobb-Douglas, 1980-2011", sep=""), digit = c(4, 4, 2, 2, 1))
  #print(tableCD)
  return(tableCD)  
}
@

@@

<<cobbDouglasEnergy function>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyColumnName is the name of the column to be used for energy.
#   This may be any of heat ("iQ"), exergy ("iX"), or useful work ("iU").
# energyString is the string to be used in table captions reprsenting the type of energy. Typically, "q", "x", or "u"
##
cobbDouglasEnergyData <- function(countryName, energyString){

  energyColumnName  <- paste("i", energyString, sep="")
  #print(energyColumnName)

@@

  summaryCDe <- summary(modelCDe) # Gives the nls summary table
  #print(summaryCDe)

@@

  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCDe <- data.frame(lambda = lambda, alpha = alpha, beta = beta, gamma = gamma)
  #print(estCDe);
  #row.names(estCDe) <- "Cobb-Douglas with e: $y = e^{\\lambda t}k^{\\alpha}l^{\\beta}e^{\\gamma}$"
  row.names(estCDe) <- paste("C-D with ", energyString, sep="")
  # The [1] subscripts pick off the lower confidence interval
  lowerCDe <- data.frame(lambda = ciCDe["lambda","2.5%"], alpha = ciCDe["a", "2.5%"], beta = betaCICDe[1], gamma = gammaCICDe[1])
  row.names(lowerCDe) <- "- 95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCDe <- data.frame(lambda = ciCDe["lambda","97.5%"], alpha = ciCDe["a", "97.5%"], beta = betaCICDe[2], gamma = gammaCICDe[2])
  row.names(upperCDe) <- "+ 95% CI"

@@

cobbDouglasEnergyTable <- function(countryName, energyType){
  dataCDe <- cobbDouglasEnergyData(countryName, energyType)
  tableCDe <- xtable(dataCDe, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
  return(tableCDe)
}

@

\subsection{Cobb-Douglas with Q}

<<CDe analysis>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDqTables <- lapply(countries, cobbDouglasEnergyTable, energyType="Q")
@

@@

\subsection{Cobb-Douglas With X}

<<cdx>>=
energyColumn <- "iX"; energyString <- "x"
dataCDx <- cobbDouglasEnergy(countryName, energyColumn, energyString)
tableCDx <- xtable(dataCDx, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
@

<<CDx Results, results='asis'>>=
print(tableCDx, caption.placement="top")
@

\subsection{Cobb-Douglas With U}

<<cdu>>=
energyColumn <- "iU"; energySTring <- "u"
dataCDu <- cobbDouglasEnergy(countryName, energyColumn, energyString)
tableCDu <- xtable(dataCDu, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
@

<<CDu Results, results='asis'>>=
print(tableCDu, caption.placement="top")
@

\section{CES With Q}

<<CES function>>=

ces <- function(countryName, energyColumnName, energyString){

@@

  # Assemble the data into data frames for the table.
  estCES <- data.frame(phi = coef(modelCES)["phi"], alpha = alpha, 
                       beta = coef(modelCES)["beta"], zeta = coef(modelCES)["zeta"], 
                       lambda_L = coef(modelCES)["lambda_L"], lambda_E = coef(modelCES)["lambda_E"])
  row.names(estCES) <- paste("CES with ", energyString, sep="")
  #print(estCES)  
  # The [1] subscripts pick off the lower confidence interval
  lowerCES <- data.frame(phi = ciCES["phi","2.5%"], alpha = alphaCICES[1], 
                         beta = ciCES["beta", "2.5%"], zeta = ciCES["zeta", "2.5%"],
                         lambda_L = ciCES["lambda_L", "2.5%"], lambda_E = ciCES["lambda_E", "2.5%"])
  row.names(lowerCES) <- "- 95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCES <- data.frame(phi = ciCES["phi","97.5%"], alpha = alphaCICES[2], 
                         beta = ciCES["beta", "97.5%"], zeta = ciCES["zeta", "97.5%"],
                         lambda_L = ciCES["lambda_L", "97.5%"], lambda_E = ciCES["lambda_E", "97.5%"])
  row.names(upperCES) <- "+ 95% CI"

@@

  #xyplot( resid(modelCESQ) ~ fitted(modelCESQ) )
  #histogram( ~resid(modelCESQ) )
  #qqmath( ~resid(modelCESQ) )
}
@

<<CES with Q>>=
countryName <- "US"
energyColumn <- "iQ"; energyString <- "q"
dataCESq <- ces(countryName, energyColumn, energyString)
tableCESq <- xtable(dataCESq, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 1, 2, 2, 6, 5, 5))
@

<<CESq Results, results='asis'>>=
print(tableCESq, caption.placement="top")
@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Caleb, put your LaTeX code here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

<<Load Data, include=FALSE>>=
#################################
# This function loads data given a country name.
# The file name from which data will be loaded is assumed to be of the form 
# "<countryName>Data.txt". The file is assumed to exist in a subfolder of this project called "data".
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a matrix with the data that has been loaded
##
loadData <- function(countryName){
  # Read the data file as a table with a header.  
  fileName <- paste("data/", countryName, "Data.txt", sep="")
  dataTable <- read.table(fileName, header = TRUE)
  # Identifies the header names associated with dataTable
  #print(names(dataTable))
  return(dataTable)
}
@

@@

#################################################
# Calculates parameter estimates and confidence intervals
# for the Cobb-Douglas production function given a country.
#
# countryName is a string containing the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a vector of data for the Cobb-Douglas model. 
# First item is the +95% CI on all parameters
# Second item contains the parameter estimates
# Third item is the -95% CI on all parameters
# Each row has names: lambda, alpha, beta, gamma, corresponding to the parameters in the model.
##
cobbDouglasData <- function(countryName){

@@

####################################
# Creates a LaTeX printable table from the Cobb Douglas data. This function first calls cobbDouglasData.
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
#
# returns a printable LaTeX table from xtable.
##
cobbDouglasTable <- function(countryName){
  dataCD <- cobbDouglasData(countryName) 
  colnames(dataCD) <- c("$\\lambda$", "$\\alpha$", "$\\beta$", "$\\gamma$")
  tableCD <- xtable(dataCD, caption=paste(countryName, " Cobb-Douglas, 1980-2011", sep=""), digit = c(4, 4, 2, 2, 1))
  #print(tableCD)
  return(tableCD)  
}
@

@@

<<cobbDouglasEnergy function>>=
###############################
# This function fits the Cobb Douglas production function including energy to historical data
# The functional form is
#
# iGDP = exp(lambda * itime) * iCapStk^alpha * iLabor^beta * iEnergy^gamma
# 
# iYear: time indexed to 0.0 at the first year [years since beginning of data set]
# iGDP = time series GDP data indexed to 1.0 at the first year [-]
# iCapStk = time series capital stock data indexed to 1.0 at the first year [-]
# iLabor = time series labor data indexed to 1.0 at the first year [-]
# iEnergy = time series energy data indexed to 1.0 at the first year [-]. 
#      This may be any of heat (iQ), exergy (iX), or useful work (iU).
# 
# returns a data object containing three rows and 5 columns.
#   col 1: lambda
#   col 2: alpha
#   col 3: beta
#   col 4: gamma
#   row 1: -95% CI for each parameter
#   row 2: estimate for each parameter
#   row 3: +95% CI for each parameter
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
##
cobbDouglasEnergyData <- function(countryName, energyType){

  print(countryName)
  energyColumnName  <- paste("i", energyType, sep="")
  #print(energyColumnName)

@@

  summaryCDe <- summary(modelCDe) # Gives the nls summary table
  print(summaryCDe)

@@

  # Combine all estimates and their confidence intervals into data frames with intelligent row names
  estCDe <- data.frame(lambda = lambda, alpha = alpha, beta = beta, gamma = gamma)
  #print(estCDe);
  #row.names(estCDe) <- "Cobb-Douglas with e: $y = e^{\\lambda t}k^{\\alpha}l^{\\beta}e^{\\gamma}$"
  row.names(estCDe) <- paste("C-D with ", energyType, sep="")
  # The [1] subscripts pick off the lower confidence interval
  lowerCDe <- data.frame(lambda = ciCDe["lambda","2.5%"], alpha = ciCDe["a", "2.5%"], beta = betaCICDe[1], gamma = gammaCICDe[1])
  row.names(lowerCDe) <- "- 95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCDe <- data.frame(lambda = ciCDe["lambda","97.5%"], alpha = ciCDe["a", "97.5%"], beta = betaCICDe[2], gamma = gammaCICDe[2])
  row.names(upperCDe) <- "+ 95% CI"

@@

####################################
# Creates a LaTeX printable table from the CES data. This function first calls cobbDouglasEnergyData.
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
#
# returns a printable LaTeX table from xtable.
##
cobbDouglasEnergyTable <- function(countryName, energyType){
  dataCDe <- cobbDouglasEnergyData(countryName, energyType)
  tableCDe <- xtable(dataCDe, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 4, 2, 2, 3))
  return(tableCDe)
}
@

\subsection{Cobb-Douglas with $Q$}

<<CDq analysis>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDqTables <- lapply(countries, cobbDouglasEnergyTable, energyType="Q")
@

@@

\subsection{Cobb-Douglas With $X$}

<<CDx analysis>>=
# Note that the anlaysis of ZA is taking a long time here. Need to figure out why.
CDxTables <- lapply(countries, cobbDouglasEnergyTable, energyType="X")
@

<<CDx Results, results='asis'>>=
print(CDxTables[["US"]], caption.placement="top")
print(CDxTables[["ZA"]], caption.placement="top")
@

\subsection{Cobb-Douglas With $U$}

<<CDu analysis>>=
CDuTables <- lapply(countries, cobbDouglasEnergyTable, energyType="U")
@

<<CDu Results, results='asis'>>=
print(CDuTables[["US"]], caption.placement="top")
print(CDuTables[["ZA"]], caption.placement="top")
@

\section{CES}

<<CES function>>=
cesData <- function(countryName, energyType){
  energyColumnName <- paste("i", energyType, sep="")
  # Load the data that we need.
  dataTable <- loadData(countryName)

@@

  # Assemble the data into data frames for the table.
  estCES <- data.frame(phi = coef(modelCES)["phi"], alpha = alpha, 
                       beta = coef(modelCES)["beta"], zeta = coef(modelCES)["zeta"], 
                       lambda_L = coef(modelCES)["lambda_L"], lambda_E = coef(modelCES)["lambda_E"])
  row.names(estCES) <- paste("CES with ", energyType, sep="")
  #print(estCES)  
  # The [1] subscripts pick off the lower confidence interval
  lowerCES <- data.frame(phi = ciCES["phi","2.5%"], alpha = alphaCICES[1], 
                         beta = ciCES["beta", "2.5%"], zeta = ciCES["zeta", "2.5%"],
                         lambda_L = ciCES["lambda_L", "2.5%"], lambda_E = ciCES["lambda_E", "2.5%"])
  row.names(lowerCES) <- "- 95% CI"
  # The [2] subscripts pick off the lower confidence interval
  upperCES <- data.frame(phi = ciCES["phi","97.5%"], alpha = alphaCICES[2], 
                         beta = ciCES["beta", "97.5%"], zeta = ciCES["zeta", "97.5%"],
                         lambda_L = ciCES["lambda_L", "97.5%"], lambda_E = ciCES["lambda_E", "97.5%"])
  row.names(upperCES) <- "+ 95% CI"

@@

  #xyplot( resid(modelCESQ) ~ fitted(modelCESQ) )
  #histogram( ~resid(modelCESQ) )
  #qqmath( ~resid(modelCESQ) )
}

####################################
# Creates a LaTeX printable table from the CES data. This function first calls cesData.
#
# countryName is a string containint the 2-letter abbreviation for the country, e.g. "US" or "CN"
# energyType is a string to be used in table captions reprsenting the type of energy. Typically, "Q", "X", or "U"
#
# returns a printable LaTeX table from xtable.
##
cesTable <- function(countryName, energyType){
  dataCESe <- cesData(countryName, energyType)
  tableCESq <- xtable(dataCESe, caption=paste(countryName, ", 1980-2011.", sep=""), digit = c(4, 1, 2, 2, 6, 5, 5))
}
@

\subsection{CES with $Q$}
<<CESq analysisQ>>=
countryName <- "US"
energyType <- "Q"
tableCESq <- cesTable(countryName, energyType)

#CESqTables <- lapply(countries, cesTable, energyType="Q")
@

<<CESq Results, results='asis'>>=
print(tableCESq, caption.placement="top")

#print(CESqTables[["US"]], caption.placement="top")
#print(CESqTables[["ZA"]], caption.placement="top")
@
